### 2014-05-20 Paper1/paper1.tex
Need to fix finalfig2 because of quirks with the s vs st. Not sure when I fixed that, but I fixed it spuriously.

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Title}
}
% Insert Author names, affiliations and corresponding author email.
\\
Justin Horowitz$^{1}$, 
James Patton$^{12\ast}$
\\
\bf{1} Bioengineering, University of Illinois at Chicago, Chicago, Illinois, United States
\\
\bf{2} Sensory Motor Performance Program, Rehabilitation Institute of Chicago, Chicago, Illinois, United States
\\
$\ast$ E-mail: pattonj@uic.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}

\section*{Introduction}
As machines interact more and more with humans, there is a growing need for a rigorous way to understand what people are trying to do, even if they become disturbed, distracted, or influenced by pathology. One can assume that a person who is familiar with a particular task will try to do the same thing each time provided all other conditions are equal. This presumes that what people do is on average what they intended to do and that their goal remains unchanged. However, if there is a disturbance, it is unclear what happens to a person's intentions because action is decoupled from intention. Lacking a means to discern that intention and changes in intention leads to challenges in human-machine interactions. Hence a comprehensive method for understanding what someone means to do (rather than what they actually do or claim to have wanted to do) as they interact within uncertain environments will prove useful in any area where a person might be thwarted from their intent such as national security, operation of machinery, athletic and music performance, rehabilitation, human augmentation, and artificial intelligence.

@@

Examination of movement intent during disturbance should allow us to test whether or not that intent is elemental or composed of subunits. Recent research has strived to understand the composition of movement intention during both infant development\cite{von1979observations} and stroke recovery\cite{rohrer2004submovements} where distinct straight-line subunits with bell-shaped speed profiles increasingly overlap and blend until, as in healthy adult reaching\cite{woodworth1899accuracy}, they become indistinguishable. Past attempts to isolate subunits from healthy adult reaching are inconclusive. Flash and Henis \cite{flash1991arm} demonstrated that  cued corrections produce recoverable subunits with a predictable shape and structure, but Liu and Todorov were able to reproduce this without using subunits \cite{liu2007evidence}. Novak et al. \cite{novak2002use} detected corrective subunits following disturbance during knob turning, but lacking access to the intent they instead relied on kinematic features of the disturbed trajectory. Here we develop a method to extract intent from disturbed movement, which will allow us to observe how intent changes during uncued replanning and whether that change is discrete or continuous.

In this paper, we employ bidirectional modeling: we convert intent into movement using a forward model so that we can examine the quality and uncertainty of converting movement to intent by intent extraction. We then explore sensitivity and failure modes using standard model analysis techniques. We also demonstrate the accuracy and utility of this on a preliminary dataset of healthy reaching movements disrupted by occasional, unpredictable disturbances. We hypothesize that this combination allows intent extraction with less uncertainty than the uncertainty naturally inherent in point-to-point reaching. We show that certain factors and measurements are more sensitive than others, and demonstrate the viability of this method for a providing a better understanding of what a person means to do when they move and are disturbed.

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}
The sections below describe the theoretical method, and then present our evaluations. First, we considered an idealized system using synthetic data that demonstrated the concept and provided an understanding of the computational process. Next we conducted an experiment on humans and the process was then evaluated on human subject data.

@@

\subsection*{Indices of variance-based sensitivity}
Sobol-distributed matrices were generated using MATLAB's sobolset() function and converted to parameter distributions (see Table~\ref{tab:parameters}) using inverse cumulative probability density. Distributions were clamped to the range of $\pm3$ standard deviations in order to avoid support for implausible parameter values. Combinations and calculations were made as prescribed by Saltelli et al. \cite{saltelli2010variance} to arrive at direct and indirect sensitivity indices.

@@

\subsection*{Apparatus}
A planar manipulandum (described in Patton and Mussa-Ivaldi \cite{patton2004robot}) was programmed to compensate and minimize any handle friction or mass. The MATLAB XPC-TARGET package (Natick, MA) was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.  Visual feedback was performed at 60 Hz using OpenGL. Closed loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was less than 8 ms, ensuring a visual delay less than one 60 Hz frame.

@@

\subsection*{Statistical Analysis}
The MATLAB statistics toolbox package (Natick, MA) was used to perform comparisons among subjects and trajectory sources using the nonparametric Kruskal-Wallis test with Tukey-Kramer post-hoc for group differences. Nonparametric testing was necessary due to floor effects from using unsigned measures.

@@

\subsection*{Human Intent Trajectories}
Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (figure 3). In both pulse timing conditions, intent remained straight line towards the target, supporting the approach.  When disturbances occurred early in reaching, intent showed some signs of correction mid-reach. When disturbances occurred late in reaching, intent tended to continue onto the target with intent changing on or near the target (figure 3 lower plots). Patterns across all disturbance timings and subjects revealed some systematic tendencies. Force-disturbed hand paths deviate from undisturbed reaching approximately 25 milliseconds after the onset of disturbance while intention remains comparable to undisturbed reaching for 100 milliseconds. While the hand returns within 500 milliseconds to undisturbed levels of error, intention is altered for an entire second (figure 4 shaded areas).

\section*{Discussion}
We sought to test the suitability and robustness of an algorithm that determines a person's intent during reaching motion, even if there disturbances. Sensitivity analysis on synthetic data revealed that the method was suitable. Response to parameter variations was more robust than human reaching, but significant misestimation could occur if the stiffness was inaccurately modelled. Examination of intended trajectories from force pulse-disturbed human reaching confirmed that these intentions remained statistically indistinguishable from undisturbed reaching for about 100 milliseconds following the onset of disturbance. As this is the first sound method for looking at intent following disturbance, we have only the undisturbed intention for comparison. If intent can change following disturbance, it can only do so after some period of time due to delay in processing sensory information and producing new motor commands. This represents a way to falsify the method: our determined intent should not diverge from the undisturbed intent of similar motions within this window of delay. As demonstrated in figure 4, the method passes this test until 100 milliseconds after the onset of disturbance. Differences after this point could be attributed to many factors including unmodeled reflexes, changes in arm stiffness, or genuine change in movement intent. In the context of the sensitivity analysis, it appears most likely that this primarily represents a change in intent and thus supports a first look at how intent changes when faced with a force disturbance.

$\lambda$ models seek to explain reaching intent as the muscles’ equilibrium trajectory. By deriving the technique in a general form, we discovered that our $x_d$ can be used to determine the muscle equilibrium of a $\lambda$ model and that they are only equivalent when $x_d$ is not changing. Even as this rejects a major claim of those modeling approaches, it explains a key experimental underpinning: Bizzi’s et al.'s \cite{bizzi1984posture} discovery of a “virtual trajectory” in deafferented monkeys with a moving equilibrium point that progressed smoothly from the initial to final positions. Even if the muscle equilibrium jumped abruptly, this would be insufficient to move the intended trajectory abruptly due to the presence of a virtual mass. This provides a striking and falsifiable prediction: human reaching intention cannot jump abruptly even if mass is wholly compensated.

Not only is intent slow to change, it does not necessarily change due to disturbance. While this work examines the time of onset of change, no statistical test could compellingly "fail to find" change in individual movements without markedly more data. Still, figure 3 shows reaches that are more compatible with intermittent control than with continuous optimal feedback control. In classic optimal control, the system is always updating the intent in response to errors \cite{todorov2002optimal}. Our results suggest a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}. Interestingly, disturbances that come later in the movement (figure 3, bottom panel) appeared to not dramatically affect some subjects' intent, suggesting that people may stubbornly “stick to their guns” when they know that the stiffness and damping characteristics will provide enough restoring torque to arrive at the target. Response latencies may be dependent on the time of the disturbance in the movement. Alternatively, control might be intermittent \cite{gawthrop2011intermittent} with limited opportunities to change intent. It remains to be seen if and how change in intent might be triggered.

@@

\textbf{While derivation and validation of intent determination was somewhat complicated, the outcome is a single filter that converts force and motion information into an estimate of intention. This filter incorporates the inversion of estimates of the actual human controller into the computation. While a large number of filters have already been proposed, we believe this to be a unique tool to be added to the arsenal of analytical methods for exploring human motor control. It has immediate applications to motor learning, neurorehabilitation, and any field interested in human-machine interaction.}

Determination of intended trajectories from force-disturbed movements allows hypothesis testing that was not previously possible. It is important to reiterate that because the human arm’s mass is almost always invertible, this approach can take \textbf{any} "candidate model" of the human arm and create a filter that estimates the intended trajectory. By using tests such as the straight line reaching test employed in this paper, one can evaluate which candidate produces the most plausible intended trajectory, supporting one model over another.  Instead of comparing two generative models to see which better approximates movement, it is now possible to perform element-by-element fitting. Mid-movement replanning is now falsifiable: do intended trajectories from disturbed motions differ from undisturbed motions? They do. Similarly, compositional claims, such as submovement decomposition, should now be directly testable. Finally, optimal feedback control models can now gain direct access to an intended trajectory to better understand and fit their cost functions.

Nevertheless, there are some drawbacks and limitations to these methods.  One limiting requirement is that accurate interface force measurement (within a few tenths of a newton) is needed. Such instrumentation is available, but must be used with great care in order to preserve the accuracy of the estimate. Our sensitivity analysis on synthetic data revealed that slight error can lead to a trajectory that accumulates a drift due to bias caused by the force term. More importantly, while modern force sensors can be highly accurate with very high signal to noise ratios, they tend to drift over time, leading to error that can grow if the device is not periodically tared. Our empirical approach used several methods to mitigate these effects, and hence the errors became small. It remains to be seen if there are methods to reduce the errors from force error, or better to eliminate their need altogether.

This study did not examine all disturbances. Although we have found and validated a potentially powerful tool for examining human intention as it changes (or fails to) in response to some disturbances, other disturbances. Even as we demonstrated the power of this tool, we couldn’t examine all force disturbances. Impulses and Gaussian noise have convenient spectral properties that provide some generality to this claim, but the arm model examined here is not a linear, time-invariant system. The eigenfunctions of this nonlinear system are not known and it is unclear which force disturbances might best expose critical vulnerabilities.

The approximation of stiffness appears to be the most critical factor in accuracy in estimating intent. While serious trial-to-trial stiffness misestimation is unlikely, co-contraction in response to disturbance might plausibly double stiffness mid-reach. Even as the model is insensitive to normal variations, halving or doubling this parameter due to contraction is plausible and has significant impact on accuracy. The limitations are readily overcome by performing this sensitivity analysis on each force environment of interest and taking measures to detect co-contraction where or when it occurs.

Our simple linear scaling of Burdet et al.'s stiffness model \cite{burdet2006stability} may not be all that is needed. More accurate approaches would provide us with too many free parameters through which to determine the outcome of the eventual extraction. Even scaling the constant and torque-dependent stiffness matrixes separately would enable rotation of the extracted trajectory. Our results point to the possibility of the impedance changing during motion, perhaps also as a function of time. The system was also shown to be sensitive to uncertainties in impedance. While this may be perceived as a shortcoming, switching from a state-varying impedance to ongoing time-varying determination of impedance may overcome these issues.

In any case, the general intent extraction approach presented here reveals the conditions needed to solve for the moving intent in an arbitrary dynamic process. Moreover, the plant model does not need to be explicitly formulated. Dynamic models that strive to understand intent are wide ranging, and can include mathematical models that range in application from national security to athletic performance. The plant (such as the actions of a crowd in a crowded square or the motions of the opposing team) may be too complex to model with simple linear transformations as we present above. However, by substituting some approximation or lookup table of input-outcome pairs, it is possible to determine in a variety of applications to obtain the intent, even when disturbed. It is not necessarily to allow a bomb to explode or the opposing team to be victorious to understand their plan behind the action.

@@

\begin{figure}[!ht]
\begin{center}
%\includegraphics[width=4in]{fig2.eps}
\end{center}
\caption{
{\bf Model sensitivity testing (from the variance-based method of Saltelli et al.\cite{saltelli2010variance}) reveals that the model is mostly sensitive to stiffness.} Direct sensitivity, in purple, shows how modeling error changes with the parameter. Indirect sensitivity, in teal, shows how modeling error from other parameters is affected by changes in the parameter. Variation in individual tests exceeds expected variance by two orders of magnitude.
}
\label{fig:sensitivity}
\end{figure}

@@

\begin{figure}[!ht]
\begin{center}
%\includegraphics[width=4in]{fig4.eps}
\end{center}
\caption{
{\bf Comparison of subjects' perpendicular errors across disturbed and undisturbed conditions reveals that intention is statistically indistinguishable from undisturbed reaching up until 100ms after the onset of movement.} Hand errors in the presence of disturbance (black dots, black band is median $\pm$ simultaneous 95\% confidence interval on the median) deviate almost immediately from undisturbed hand errors (green dots, band). Movement intention (red dots, band) deviates at approximately 100ms after the onset of disturbance. The hand rejoins the intention approximately 500ms after the onset of disturbance.
}
\label{fig:grouptrends}
\end{figure}

@@

\begin{table}[!ht]
\caption{
\bf{Arm Parameter Values and Sensitivities}}
\begin{tabular}{|c|c|c c|c c|}
\hline
Parameter Name &
Units &
Nominal &
SD &
$\frac{S}{\text{Total}}$ &
$\frac{ST}{\text{Total}}$ \\ \hline
Upper Arm Length ($L_1$) &
m &
0.33 &
0.01 &
0.07 &
0 \\
Forearm Length ($L_2$) &
m &
0.34 &
0.01 &
0.02 &
0 \\
Upper Arm Center of Mass Ratio ($\frac{L_{m1}}{L_1}$) &
1 &
0.436 &
0.0695 &
0.01 &
0 \\
Forearm Center of Mass Ratio ($\frac{L_{m2}}{L_2}$) &
1 &
0.682 &
0.0431 &
0.02 &
0 \\
Gross Body Mass ($m_g$) &
kg &
79.36 &
3.1 &
0.01 &
0 \\
Upper Arm Mass Ratio ($\frac{m_1}{m_g}$) &
1 &
0.028 &
0.0029 &
0.01 &
0 \\
Forearm Mass Ratio ($\frac{m_2}{m_g}$) &
1 &
0.022 &
0.0025 &
0.04 &
0 \\
Upper Arm Radius of Gyration Ratio &
1 &
0.322 &
0.0161 &
0.01 &
0 \\
Forearm Radius of Gyration Ratio &
1 &
0.468 &
0.0234 &
0.01 &
0 \\
Shoulder Parallel Coordinate &
m &
-0.057 &
0.02 &
0.13 &
0.01 \\
Shoulder Perpendicular Coordinate &
m &
0.88 &
0.02 &
0.05 &
0 \\
Force Sensor Miscalibration $x\text{-axis}$ &
N &
0 &
0.231 &
0.06 &
0 \\
Force Sensor Miscalibration $y\text{-axis}$ &
N &
0 &
0.1067 &
0.06 &
0.02 \\
Force Sensor Gaussian Noise SD $x\text{-axis}$ &
N &
0 &
0.1653 &
0.01 &
0 \\
Force Sensor Gaussian Noise SD $y\text{-axis}$ &
N &
0 &
0.2869 &
0.01 &
0 \\
Torque-Invariant Impedance Misestimation Ratio &
1 &
1 &
0.15 &
0.22 &
0.04 \\
Torque-Varying Impedance Misestimation Ratio &
1 &
1 &
0.15 &
0.11 &
0.01 \\
Damping-to-Stiffness Ratio ($k_d$) &
$\text{sec}^{-1}$ &
0.0833 &
0.0125 &
0.06 &
0 \\
Reflex Impedance Scale Factor &
1 &
0.02 &
0.003 &
0.03 &
0 \\
Reflex Damping to Stiffness Ratio ($g_d$) &
$\text{sec}^{-1}$ &
2 &
0.3 &
0.03 &
0 \\ \hline
\end{tabular}
\begin{flushleft}Synthetic model parameters and their associated mean (nominal) values and standard deviations (SD) used to determine the sensitivity indices of Saltelli et al. \cite{saltelli2010variance}. Also shown are the resulting sensitivity indices. Sensitivity indices less than one-thousandth of the total sensitivity were reported as zero. Total sensitivity was $1.8 \mu m$, showing very small average deviations when using this approach.
\end{flushleft}
\label{tab:parameters}
\end{table}

>>>>

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Intention can be Determined Despite Disturbances}
}
% Insert Author names, affiliations and corresponding author email.
\\
Justin Horowitz$^{1}$, 
James Patton$^{12\ast}$
\\
\bf{1} Bioengineering, University of Illinois at Chicago, Chicago, Illinois, United States
\\
\bf{2} Sensory Motor Performance Program, Rehabilitation Institute of Chicago, Chicago, Illinois, United States
\\
$\ast$ E-mail: pattonj@uic.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
We derive a novel modeling approach for determining movement intent from a force-disturbed planar arm trajectory, examine its parameter sensitivity, and test it on human reaching data. While we demonstrate that popular modeling approaches can be algebraically inverted to yield the same result, this more general approach reveals strengths, limitations, and underlying assumptions. After populating our model with measured lengths and masses as well as impedance values from the literature, we perform sensitivity analysis to determine the accuracy of the model relative to the uncertainty naturally inherent in point-to-point reaching. We find that sensitivity to parameter inaccuracy is on the order of microns and lower than the inaccuracy of human reaching. Further, most of the sensitivity results from uncertainty with regard to the stiffness of the joints. The robustness of this approach indicated that it was suitable for determining reaching intent from human point-to-point reaching movements disturbed by large force pulses. Intended trajectory showed no change from undisturbed reaching for 150 ms after the onset of disturbance. Beyond this point, stiffening of the arm, change in intent, or both usually lead to deviation from undisturbed reaching. Not all disturbed reaches showed evidence of change in intent. Unlike $\lambda$ models, the reaching intent determined by our model is physiologically plausible. This success allows real-time determination of error that can free learning enhancement and treatment methods like error augmentation and challenge point from needing to prescribe explicit task goals. In addition, extraction from recorded human motion enables direct inspection of and hypothesis testing on the contents of human motor intent. While not explicitly examined, this approach is also valid for determination intent is more complex situations like team sports and national security.

\section*{Introduction}
As machines interact more and more with humans, there is a growing need for a rigorous way to understand what people are trying to do, even if they become disturbed, distracted, or influenced by pathology. One can assume that a person who is familiar with a particular task will try to do the same thing each time provided all other conditions are equal. This presumes that what people do is on average what they intended to do and that their goal remains unchanged. However, if there is a disturbance, it is unclear what happens to a person's intentions because action is decoupled from intention. Challenges arise in human-machine interactions because intention and changes in intention cannot be easily discerned. Hence a comprehensive method for understanding what someone means to do (rather than what they actually do or claim to have wanted to do) as they interact within uncertain environments will prove useful in any area where a person might be thwarted from their intent such as national security, operation of machinery, athletic and music performance, rehabilitation, human augmentation, and artificial intelligence.

@@

Here we develop and analyze a method to extract intent from disturbed movement, which will allow us to observe if and how intent changes following disturbance. We employ bidirectional modeling: we convert intent into movement using a forward model so that we can examine the quality and uncertainty of converting movement to intent by intent extraction. We then explore sensitivity and failure modes using standard model analysis techniques. We also demonstrate the accuracy and utility of this on a preliminary dataset of healthy reaching movements disrupted by occasional, unpredictable disturbances. We hypothesize that this combination allows intent extraction with less uncertainty than the uncertainty naturally inherent in point-to-point reaching. We show that certain factors and measurements are more sensitive than others, and demonstrate the viability of this method for a providing a better understanding of what a person means to do when they move and are disturbed.

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Methods}
The sections below describe the theoretical method, and then present our evaluations. First, we considered an idealized system using synthetic data that demonstrated the concept and provided an understanding of the computational process. Next we conducted an experiment on humans and the process was then evaluated on human subject data.

@@

\subsection*{Indices of variance-based sensitivity}
Sobol-distributed matrices were generated using MATLAB's sobolset() function and converted to parameter distributions (see Table~\ref{tab:parameters}) using inverse cumulative probability density. Distributions were clamped to the range of $\pm3$ standard deviations in order to avoid support for implausible parameter values. Combinations and calculations were made as prescribed by Saltelli et al. \cite{saltelli2010variance} to arrive at direct and total sensitivity indices.

@@

\subsection*{Apparatus}
A planar manipulandum (described in Patton and Mussa-Ivaldi \cite{patton2004robot}) was programmed to compensate and minimize any handle friction or mass. The MATLAB XPC-TARGET package \cite{MATLAB:2008} was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.  Visual feedback was performed at 60 Hz using OpenGL. Closed loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was less than 8 ms, ensuring a visual delay less than one 60 Hz frame.

@@

\subsection*{Statistical Analysis}
The MATLAB statistics toolbox package \cite{MATLAB:2008} was used to perform comparisons among subjects and trajectory sources using the nonparametric Kruskal-Wallis test with Tukey-Kramer post-hoc for group differences. Nonparametric testing was necessary due to floor effects from using unsigned measures.

@@

\subsection*{Human Intent Trajectories}
Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (figure 3). In both pulse timing conditions, intent remained straight line towards the target, supporting the approach.  When disturbances occurred early in reaching, intent showed some signs of correction mid-reach. When disturbances occurred late in reaching, intent continued onto the target (figure 3 lower plots). We observed a systematic change of intent 150 milliseconds after the onset of force disturbance. While the hand returns within 600 milliseconds to undisturbed levels of error, intention is altered for almost 900 milliseconds (figure 4 shaded areas).

As this is the first sound method for looking at intent following disturbance, we have only the undisturbed intention for comparison. If intent can change following disturbance, it can only do so after some period of time due to delay in processing sensory information and producing new motor commands. This represents a way to falsify the method: our determined intent should not diverge from the undisturbed intent of similar motions within this window of delay. As demonstrated in figure 4, the method passes this test until 150 milliseconds after the onset of disturbance. Further

\section*{Discussion}
We sought to test the suitability and robustness of an algorithm that determines a person's intent during reaching motion, even if there disturbances. Sensitivity analysis on synthetic data revealed that the method was suitable in that errors in response to parameter variations were smaller than trial-to-trial variance observed in human reaching. Significant misestimation could occur if the stiffness was inaccurately modelled. We tested this on human reaching and found that in spite of the hand being distubed by forces, the intended movement remained on-course to the target until 120-150 ms before responding. This represents a new and accurate method for viewing intent and how it changes when faced with a force disturbance.

While the derivation of intent determination required many steps, the outcome is a filter whose use is straight-forward. Many filters for converting assumed intent into simulated hand trajectories have been proposed and validated. Demonstrating the validity and plausibility of inverting those filters to determine intent from hand trajectory adds a novel method to the arsenal of analytical methods for exploring human motor control.

This approach explains the discovery by Bizzi et al. \cite{bizzi1984posture} of a ``virtual trajectory'' ($\alpha$) in deafferented monkeys that progressed smoothly from the initial to final positions and how that virtual trajectory is different from $\lambda$. By deriving the technique in a general form, we discovered that our $x_d$ can be used to determine the muscle equilibrium of a $\lambda$ model and that they are only equivalent when $x_d$ is not changing.  Even if the muscle equilibrium jumped abruptly, this would be insufficient to move the intended trajectory abruptly due to the presence of a virtual mass. This provides a striking and falsifiable prediction: reaching intention cannot jump abruptly.

Not only is intent slow to change, it does not necessarily change following disturbance. While this work examines the time of onset of change, statistical testing cannot be used to detect the absence of change in an individual movement. Still, figure 3 shows reaches that are more compatible with intermittent control than with continuous optimal feedback control. In classic optimal control, the system is always updating the intent in response to errors \cite{todorov2002optimal}. Our results suggest a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}. Interestingly, disturbances that come later in the movement (figure 3, bottom panel) appeared to not dramatically affect some subjects' intent, suggesting that people may stubbornly ``stick to their guns'' when they know that the stiffness and damping characteristics will provide enough restoring torque to arrive at the target. Time of response may be dependent on the time of the disturbance in the movement. Alternatively, control might be intermittent \cite{gawthrop2011intermittent} with limited opportunities to change intent. It remains to be seen if and how change in intent might be triggered.

@@

Determination of intended trajectories from force-disturbed movements allows hypothesis testing that was not previously possible. It is important to reiterate that because the human arm’s mass is almost always invertible, this approach can take \textbf{any} ``candidate model'' of the human arm and create a filter that estimates the intended trajectory. By using tests such as the comparison with undisturbed straight line reaching employed in this paper, one can evaluate which candidate produces the most plausible intended trajectory, supporting one model over another.  Instead of comparing two generative models to see which better approximates movement, it is now possible to perform element-by-element fitting. Mid-movement replanning is now falsifiable: do intended trajectories from disturbed motions differ from undisturbed motions? They do. Similarly, compositional claims, such as whether movements are comprised of subunits, should now be directly testable. Finally, optimal feedback control models can now gain direct access to an intended trajectory to better understand and fit hypothesized cost functions.

The plant model does not need to be explicitly formulated. Dynamic models that strive to understand intent are wide ranging, and can include mathematical models that range in application from national security to athletic performance. The plant (such as the actions of a crowd in a crowded square, the motions of the opposing team,) may be too complex to model with simple linear transformations as we present above. However, by substituting some approximation or lookup table of input-outcome pairs, it is possible in a variety of applications to obtain the intent, even when disturbed. It is not necessarily to allow a bomb to explode or the opposing team to be victorious to understand the plan behind their action.

Nevertheless, there are some limitations to these methods.  One limiting requirement is that accurate interface force measurement (within a few tenths of a newton) is needed. Such instrumentation is available, but must be used with great care in order to preserve the accuracy of the estimate. Our sensitivity analysis on synthetic data revealed that slight error can lead to a trajectory that accumulates a drift due to bias caused by the force term. More importantly, while modern force sensors can be highly accurate with very high signal to noise ratios, they tend to drift over time, leading to error that can grow if the device is not periodically tared. Our empirical approach used several methods to mitigate these effects, and hence the errors became small. It remains to be seen if there are methods to further reduce the misestimation due to force error or eliminate the need for force estimation altogether.

The approximation of stiffness appears to be the most critical factor in accuracy in estimating intent. While serious trial-to-trial stiffness misestimation is unlikely, co-contraction in response to disturbance might plausibly double stiffness mid-reach. Even as the model is insensitive to normal variations, a human can volitionally halve or double their stiffness \cite{franklin2007endpoint}. Co-contraction is plausible and has significant impact on accuracy. These limitations are readily overcome by performing this sensitivity analysis on each force environment of interest and taking measures to detect co-contraction and hence stiffness alterations. Additionally, real-time determination of intent may allow real-time determination of stiffness.

Our simple linear scaling of Burdet et al.'s stiffness model \cite{burdet2006stability} facilitated hypothesis testing, but was inherently limiting. More accurate approaches would provide us with too many free parameters through which to determine the outcome of the eventual extraction and obtain a spurious favorable result. Our results point to the possibility of the impedance changing during motion, perhaps as a function of time. As mentioned above, while this may be perceived as a shortcoming, it was necessary and potentially facilitates real-time determination of stiffness in future work.

The general intent extraction approach presented here reveals the conditions needed to solve for the moving intent in an arbitrary dynamic process.

@@

\begin{figure}[!ht]
\begin{center}
%\includegraphics[width=4in]{fig2.eps}
\end{center}
\caption{
{\bf Model sensitivity testing (from the variance-based method of Saltelli et al.\cite{saltelli2010variance}) reveals that the model is mostly sensitive to stiffness.} Parameter values are varied in a Sobol-distributed fashion according to Table 1. First order sensitivity, in purple, is the error that would be removed if the parameter was fixed at its nominal value. Total sensitivity, in teal, shows error that would remain if all other parameters were fixed at their nominal values.
}
\label{fig:sensitivity}
\end{figure}

@@

\begin{figure}[!ht]
\begin{center}
%\includegraphics[width=4in]{fig4.eps}
\end{center}
\caption{
{\bf Comparison of subjects' perpendicular errors across disturbed and undisturbed conditions reveals that intention is statistically indistinguishable from undisturbed reaching up until 100ms after the onset of movement.} Hand errors in the presence of disturbance (black dots, black band is median $\pm$ simultaneous 95\% confidence interval on the median) begin to deviate from undisturbed hand errors (green dots, band) beginning 95 ms after the onset of disturbance. Movement intention (red dots, band) deviates at 150ms. The hand rejoins the intention approximately 500ms after the onset of disturbance.
}
\label{fig:grouptrends}
\end{figure}

@@

\begin{table}[!ht]
\caption{
\bf{Arm Parameter Values and Sensitivities}}
\begin{tabular}{|c|c|c c|c c|}
\hline
Parameter Name &
Units &
Nominal &
SD &
$S$ &
$S_T$ \\ \hline
Upper Arm Length ($L_1$) &
m &
0.33 &
0.01 &
0 &
0.10 \\
Forearm Length ($L_2$) &
m &
0.34 &
0.01 &
0 &
0.03 \\
Upper Arm Center of Mass Ratio ($\frac{L_{m1}}{L_1}$) &
1 &
0.436 &
0.0695 &
0 &
0.02 \\
Forearm Center of Mass Ratio ($\frac{L_{m2}}{L_2}$) &
1 &
0.682 &
0.0431 &
0 &
0.02 \\
Gross Body Mass ($m_g$) &
kg &
79.36 &
3.1 &
0 &
0.02 \\
Upper Arm Mass Ratio ($\frac{m_1}{m_g}$) &
1 &
0.028 &
0.0029 &
0 &
0.01 \\
Forearm Mass Ratio ($\frac{m_2}{m_g}$) &
1 &
0.022 &
0.0025 &
0 &
0.06 \\
Upper Arm Radius of Gyration Ratio &
1 &
0.322 &
0.0161 &
0 &
0.01 \\
Forearm Radius of Gyration Ratio &
1 &
0.468 &
0.0234 &
0 &
0.01 \\
Shoulder Parallel Coordinate &
m &
-0.057 &
0.02 &
0.02 &
0.18 \\
Shoulder Perpendicular Coordinate &
m &
0.88 &
0.02 &
0 &
0.08 \\
Force Sensor Miscalibration $x\text{-axis}$ &
N &
0 &
0.231 &
0 &
0.08 \\
Force Sensor Miscalibration $y\text{-axis}$ &
N &
0 &
0.1067 &
0.03 &
0.08 \\
Force Sensor Gaussian Noise SD $x\text{-axis}$ &
N &
0 &
0.1653 &
0 &
0.01 \\
Force Sensor Gaussian Noise SD $y\text{-axis}$ &
N &
0 &
0.2869 &
0 &
0.01 \\
Torque-Invariant Impedance Misestimation Ratio &
1 &
1 &
0.15 &
0.06 &
0.31 \\
Torque-Varying Impedance Misestimation Ratio &
1 &
1 &
0.15 &
0.0 &
0.15 \\
Damping-to-Stiffness Ratio ($k_d$) &
$\text{sec}^{-1}$ &
0.0833 &
0.0125 &
0 &
0.08 \\
Reflex Impedance Scale Factor &
1 &
0.02 &
0.003 &
0 &
0.04 \\
Reflex Damping to Stiffness Ratio ($g_d$) &
$\text{sec}^{-1}$ &
2 &
0.3 &
0 &
0.04 \\ \hline
\end{tabular}
\begin{flushleft}Synthetic model parameters and their associated mean (nominal) values and standard deviations (SD) used to determine the sensitivity indices of Saltelli et al. \cite{saltelli2010variance}. Also shown are the resulting sensitivity indices. Sensitivity indices less than one-thousandth of the total sensitivity were reported as zero. Total sensitivity was $1.3 \mu m^2$, showing very small average deviations when using this approach. First order sensitivity, $S$, can be interpreted as fraction of variance removable by perfectly correcting a factor while total sensitivity, $S_T$ can be interpreted as the fraction left after correcting all other factors.
\end{flushleft}
\label{tab:parameters}
\end{table}
